\documentclass[a4paper,twoside=false,abstract=false,numbers=noenddot,
titlepage=false,headings=small,parskip=half,version=last]{scrartcl}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}

\usepackage[colorlinks=true, pdfstartview=FitV,
linkcolor=black, citecolor=black, urlcolor=blue]{hyperref}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{multirow}

\usepackage{tikz}
\usetikzlibrary{matrix}

\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}

\usepackage{fullpage}

\theoremstyle{definition}
\newtheorem{exercise}{Exercise}

\theoremstyle{remark}
\newtheorem*{solution}{Solution}
\newtheorem*{remark}{Remark}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\newcommand{\NN}{\ensuremath{\mathbb{N}}}
\newcommand{\ZZ}{\ensuremath{\mathbb{Z}}}
\newcommand{\QQ}{\ensuremath{\mathbb{Q}}}
\newcommand{\RR}{\ensuremath{\mathbb{R}}}
\newcommand{\CC}{\ensuremath{\mathbb{C}}}
\newcommand{\Fourier}{\ensuremath{\mathcal{F}}}
\newcommand{\Laplace}{\ensuremath{\mathcal{L}}}


\DeclareMathOperator{\Hom}{Hom}
\DeclareMathOperator{\End}{End}
\DeclareMathOperator{\im}{im}
\DeclareMathOperator{\id}{id}


\renewcommand{\labelenumi}{(\alph{enumi})}

\author{Jim Holmström - 890503-7571}
\title{Avancerade algoritmer DD2440}
\subtitle{Homework D}

\begin{document}

\maketitle
\thispagestyle{empty}

%-----------------------
\begin{exercise}
{\bf
Analyze the amortized complexity of a series of inserts and deletes assuming.
\begin{itemize}
    \item The table starts empty (num=0) and with (size = 0) 
    \item If the table is of size zero and an element is added it grows to size 1.
    \item If a tabl of size $s \le 1$ is full and an element is added, the table is grown to size $2n$, elements are relocated and the new element is added.
    \item If, after removal, a table of size $s \le 1$ has less than $s/3$ elements is shrunk to size $\lceil 2s/3 \rceil$, and the elements are relocated.
    \item The cost of adding or deleting an element is 1 (except if the table grows or shrinks)
    \item The cost of growing or shrinking the table is equal to the number of elements moved as a result of relocation.
\end{itemize}
Use the potential function $\Phi = |2$num$-$size$|$
}
\end{exercise}
\begin{solution}
Define: $alpha_i = \frac{num_i}{size_i}$

We always know that $ 1/3 < \\alpha_{i-1} < 1$ else the size would have already
increased or decreased.
$\Phi_i = |2num_i-size_i|=|(2\alpha_i-1)size_i|=|c_i|$

\begin{equation}
    \[ 
   |x| = \left\{\begin{array}{ll}
                    x & $x \geq 0$\\
                   -x & $x < 0$.
         \end{array}\right. 
    \] 
\end{equation}
\begin{lemma}
\label{absfunction}
Using the fact that $b>a,|c-a|-|c-b|$ is $(b-a)$ when $c \le a$, $(a-b)$ when
$c \ge b$ and takes every value (in the continuous case) in between when
$a<c<b$ by continuity reasons, more precisely $2c-(a+b)$  
\end{lemma}
Insert:\\

Without expand we have $size_i=size_{i-1} \wedge num_i=num_{i-1}+1$

\begin{equation}
\begin{split}
    \hat{c}_i & = c_i + \Phi_i - \Phi_{i-1} \\
              & = 1 + |2num_i-size_i|-|2num_{i-1}-size_{i-1}| \\
              & = 1 + |c_i| - |c_i-2| \\
              & = 1 + \[\left\{\begin{array}{ll}
                            c_i-c_i-2 & =-2 & c_i\ge 2\\
                           -c_i+c_i+2 & = 2 & c_i\le 0\\
                                      &   0 & c_i  = 1
                        \end{array}\right. \]  
\end{split}
\end{equation}

With expand we have $size_i=2\cdot size_{i-1} \wedge num_i=num_{i-1}+1$

\begin{equation}
\begin{split}
    \hat{c}_i & = c_i + \Phi_i - \Phi_{i-1} \\
              & = 1 + |2num_i-size_i|-|2num_{i-1}-size_{i-1}| \\
              & = 1 + |2(num_{i-1}+1)-size_{i-1}/2| - |2(num_{i-1}-size_{i-1}| \\
              & = 1 + |2(\alpha_{i-1}-1/2)size_{i-1}+2| - |(2\alpha_{i-1}-1)size_{i-1}| \\
              & = 1 + (2\alpha_{i-1}-1/2-|2\alpha_{i-1}-1|)size_{i-1} + 2 \\
              &  \{\mbox{Can for all ranges find a $\alpha$ that eliminates
                     $size_{i-1}$ and bounds the equation}\} \\
              & < 0*size_{i-1} + 3 = 3

\end{split}
\end{equation}

Delete:\\

%$\alpha_{i-1}$

%In the case $\alpha_i \ge 1/2$.\\
Without shrinking \\
\begin{equation}
\begin{split}
    \hat{c}_i & = c_i + \Phi_i - \Phi_{i-1} \\
              & = 1 + |2num_i-size_i|-|2num_{i-1}-size_{i-1}| \\
              & = 1 +  
\end{split}
\end{equation}

%In the case $\alpha < 1/2$.\\
With shrinking \\
\begin{equation}
\begin{split}
    \hat{c}_i & = c_i + \Phi_i - \Phi_{i-1} \\
              & = 1 + |2num_i-size_i|-|2num_{i-1}-size_{i-1}| \\
              & =  
\end{split}
\end{equation}




And since bounded by a constant the amortized cost for delete is $O(1)$
That is any sequence of $n$ operations is $O(n)$

\end{solution}
%\clearpage
%-----------------------
\begin{exercise}
{\bf
Let \verb+makeSet+,\verb+union+, and \verb+findSet+ be the three operations of the union-find data structore implmented using a disjoint-set forest with union by rank and path compression.
Draw a picture of the data structure that results from running the following code. 
\begin{verbatim}
for (i = 1; i <= 16; i++)
    makeSet(i)
for (i = 1; i <= 15; i += 2)
    union(i, i+1)
for (i = 5; i <= 13; i += 4)
    union(i, i+2)
#draw the datastructure at this point in time
union(3, 15)
union(10, 8)
union(16, 2)
union(13, 9)
#draw the datastructure at this point in time
\end{verbatim}
}
\end{exercise}
\begin{solution}
\begin{verbatim}
(value,rank,p)
\end{verbatim}
After first for-loop:\\
\begin{verbatim}
( 1, 0, 2)
( 2, 1, 2)

( 3, 0, 4)
( 4, 1, 4)

( 5, 0, 6)
( 6, 1, 6)

( 7, 0, 8)
( 8, 1, 8)

( 9, 0,10)
(10, 1,10)

(11, 0,12)
(12, 1,12)

(13, 0,14)
(14, 1,14)

(15, 0,16)
(16, 1,16)
\end{verbatim}
After second for-loop:\\

\begin{verbatim}
( 1, 0, 2)
( 2, 1, 2)

( 3, 0, 4)
( 4, 1, 4)

( 5, 0, 6)
( 6, 1, 8)
( 7, 0, 8)
( 8, 2, 8)

( 9, 0,10)
(10, 1,12)
(11, 0,12)
(12, 2,12)

(13, 0,14)
(14, 1,16)
(15, 0,16)
(16, 2,16)
\end{verbatim}
After the three first unions:\\
\begin{verbatim}
( 1, 0, 2)
( 2, 2, 2) //union(16,2)

( 3, 0, 4)
( 4, 1,16) //union(3,15)

( 5, 0, 6)
( 6, 1, 8)
( 7, 0, 8)
( 8, 3, 8) //union(10,8)

( 9, 0,10)
(10, 1,12)
(11, 0,12)
(12, 2, 8) //union(10,8)

(13, 0,14)
(14, 1,16)
(15, 0,16)
(16, 2,16)

\end{verbatim}
After the last union:\\
\begin{verbatim}
( 1, 0, 2)
( 2, 2, 2) 

( 3, 0, 4)
( 4, 1,16)

( 5, 0, 6)
( 6, 1, 8)
( 7, 0, 8)
( 8, 3, 8)

( 9, 0, 8) //2nd path compression
(10, 1, 8) //3rd path compression
(11, 0,12)
(12, 2, 8)

(13, 0,16) //1st path compression
(14, 1,16)
(15, 0,16)
(16, 2, 8) //union(13,9)
\end{verbatim}
\end{solution}
%\clearpage
%-----------------------
\begin{exercise}
{\bf
In class we analyzed disjoint-set forest with union by rank with path compression.
In the analsysis we used without proof that for any element, its rank never
exceeds $log_2n$, prove this. More precisly.
Show that in a sequence of $m$ operation where $n$ operations are
\verb+makeSet+ no element gets rank more than $log_2n$.
}
\end{exercise}
\begin{solution}

Since there is no way to connect something that doesn't already have a
connection (except for starting one with constant length) else there will
always be path-compression. <...> lite luddigt men iden kommer fram iaf

Having multiple trees will trivially only increase the number of components for
a given rank/depth.

Linking together two trees will always result in at least two links to the root
and when doing path compression

Path compression just adds to the compression of the treedepth and one way to
trying to avoid path compression is to just JOIN(root_i,root_j)
and this will always (except for the start case when x.p.p=x.p but this will
only happen when tree is of depth =2, is this true? Verify) yield a double
connection. And now since we know that all connections except for a constant
amount near the leaves have atleast 2 connections (more with pathcompression)
when can easily conclude that the one disjoint set will have atleast 2^k (<---
must include the k value that is the rank, BALANCE IS KEY, the smaller tree
will always come under the ) number
path compression must be included!!!

rank will only increase when the to joined trees are of the same length


of elements

\end{solution}
%\clearpage
%-----------------------
\begin{exercise}
{\bf
There are $n$ professors and $2n$ classrooms. The input is a table \verb+t[1..n,1..2*n]+ where \verb+t[p, c]+ is the distance from the office of professor p to a classroom c.
Design an efficient algorithm that finds a way to assign each professor to a classroom so that each professor gets a classroom and no classroom is used by more than one professor, and maximize the sum of the distance the professors need to walk if they all walk from their offices to their classrooms.
Your algorithm should run in time polynomial in $n$, and you may analyze it using unit cost.
For full credit, its complexity should be noticeably better then $O(n^4)$
}
\end{exercise}
\begin{solution}
Firstly create $2n-n=n$ ``empty-classroom''-professors with all their distances to the classrooms set to zero $(n^2)$ this is to make a perfect matching.
Then you want to make an assignment of professors,including the ``empty-classroom''-professors to classrooms, so that each professor gets an unique classroom.
The problem can then be easily solved as a maximum weighted matching problem in a bipartite graph by for example the Hungarian algorihm running in $O(n^3)$

This results in a solution in the running time $O(n^3) \qed$
\end{solution}
%\clearpage
%-----------------------
\begin{exercise}
{\bf
This problem is the same as Problem 4, except that we have a different target function.
This time you should maximize the distance that the professor with the shortest path need to take.
}
\end{exercise}
\begin{solution}

Sort all edges according to their weights in a new list with (value,edgeindex). $O(|E|+|E|log(|E|))=O(n^2log(n))$

Binarysearch $w_i$ in the sorted edgelist: //$O(log(|E|)=log(n))$
    \verb+first+ = get the first edge with this value $O(|E|)=O(n^2)$
    Create a new unweighted set of edges $uw_i$ from \verb+first+ and higher that have weight $w_i$ or
    higher. $O(|E|)=O(n^2)$
    if(bipartitmatching($uw_i$)=n): //it exists a "perfect" match
        try higher value in the binarysearch
    else:
        try lower value in the binarysearch 

Bipartite-matching:
   ...     
    
\end{solution}

\end{document}
