The TSP can be divided into two parts, constructing an initial path and
optimizing the constructed path by local optimization.
The section \ref{sec:constructing_the_path}
discusses various heuristics used for constructing the path and section
\ref{sec:optimizing_the_path} describes heuristics for optimzing that path.  

For all the heuristics, it is easier to view the TSP as a graph problem.
The TSP can be viewed as a graph problem, where each point $(x, y)$ is a node in
the graph, and there is an undirected edge from each point to all other points. 
The weight of each edge is the distance between the two end-points of the edge.
The TSP then reduces to find the minimum hamiltonian cycle.

\subsection{Constructing the path}
\label{sec:constructing_the_path}
This section will describe two different heuristics for constructing an
initial path.

\subsubsection{Nearest neighbour}
\label{sec:nearest_neighbour}
\paragraph{Heuristic}
\begin{codebox}
\Procname{$\proc{Nearest-Neighbour}(\id{cities})$}
\li $\id{visited} \gets \proc{Array}(n)$
\li \For $i \gets 0$ \To $n$
\li     \Do
            $\id{visited}[i] \gets \const{false}$
        \End
\li $\id{path} \gets \proc{Vector}()$
\li $\id{city} \gets 0$
\li \While $\proc{Size}(visited) < \proc{Size}(\id{cities})$
\label{li:nn:outer}
\li     \Do
            $\id{visited}[city] \gets \const{true}$
\li         $\proc{Append}(\id{path}, \id{city})$
\li         $min \gets \infty$
\li         $next \gets -1$
\li         \For $i \gets 0$ \To $\proc{Size}(cities)$ \label{li:nn:inner}
\li             \Do
                    \If $\proc{Distance}(\id{city}, i) < min$ and
                        $\id{visited}[i] \isequal \const{false}$
\li                     \Then
                            $min \gets \proc{Distance}(\id{city}, i)$
\li                         $next \gets i$
                        \End
                \End
\li         $city \gets next$
        \End
\li \Return $\id{path}$
\end{codebox}
\paragraph{Description} \proc{Nearest-Neighbour} simply starts at the first 
city, then repeatedly selects the closes city until all cities have been
visited. To keep track of which of the cities that have been visited,
\proc{Nearest-Neighbour} stores all the visited cities in a set.
\paragraph{Running time}
Let $n$ be the number of cities to visit. The while loop on line
\ref{li:nn:outer} will then run $n$ times, since one city is added to set
\id{visited} each time. The loop on line \ref{li:nn:inner} will run $n$ times,
since it iterates over all the cities. The \proc{Distance} operation have
running time $O(1)$, so the running time of the inner loops body is $O(1)$.
Therefore, the running time of the outer loops body is $O(n)$ and the total
running time is $O(n^2)$.
\paragraph{Length of path}
Let $n$ be the number of cities, $opt$ be the optimal path for visiting the $n$
cities and $nn$ be the path
found by \proc{Nearest-Neighbour}. The relation between $opt$ and $nn$ for any instance $I$ is then
\[\frac{nn(I)}{opt(I)} \leq \frac{1}{2}(\lceil \log_2(n) \rceil + 1)\]
according to~\cite{johnson}. This means that for all instances, the path found
by \proc{Nearest-Neighbour} will be at most 
$\frac{1}{2}(\lceil \log_2(n) \rceil + 1)$ larger than the optimal path.

\clearpage

\subsubsection{Greedy}
\paragraph{Heuristic}
\begin{codebox}
\Procname{$\proc{Greedy}(cities)$}
\li $n \gets \proc{Size}(cities)$
\li $deg = \proc{Array}(n, 0)$ \Comment Every element initialized to 0
\li $h = \proc{Min-Heap}()$
\li \For $i \gets 0$ \To $n$
\li     \Do
            \For $j \gets 0$ \To $n$
\li             \Do
                    $d \gets \proc{Distance}(i,j)$
\li                 $\proc{Insert}(h, (d, i, j))$ \Comment Sorted by $d$
                \End
        \End
\li $edges \gets \proc{Vector}()$
\li \While $\proc{Size}(edges) < n$ \label{li:greedy:while}
\li     \Do
            $(d, i, j) = \proc{Heap-Extract-Min}(h)$
\li         \If $deg[i] < 2$ and $((\proc{Size}(edges) \isequal (n-1)$ and
                $deg[j] \isequal 1)$ or $deg[j] \isequal 0)$
\li             \Then
                    $\proc{Append}(edges, (i, j))$
\li                 $deg[i] \gets deg[i] + 1$
\li                 $\id{in-path}[i] \gets \const{true}$
\li                 $deg[j] \gets deg[j] + 1$
\li                 $\id{in-path}[j] \gets \const{true}$
\li             \Else \textbf{continue}
                \End
        \End
\li $path \gets \proc{Vector}()$
\li $last \gets edges[0]$
\li $\proc{Append}(path, \attrib{last}{to})$
\li \While $\proc{Size}(path) < n$
\li     \Do
            \For $e \in edges$
\li             \Do
                    \If $\attrib{last}{to} \isequal \attrib{e}{from}$
\li                     \Then
                            $\proc{Append}(path, \attrib{e}{to})$
\li                         $last \gets e$
\li                         \textbf{break}
                        \End
                \End
        \End
\li \Return $path$
\end{codebox}
\paragraph{Description} \proc{Greedy} is very similar to
\proc{Nearest-Neighbour}, but instead of choosing the closest city to move to,
\proc{Greedy} always chooses the shortest edge in the graph (under some
constraints). In order for \proc{Greedy} to form a complete tour of all the
edges, the edges has to be chosen in a certain way. When a new edge is added to
the tour, the degree of the nodes at the ends of the edge must have a degree of
either 0 or 1. This is because a tour is only allowed to visit each city
\emph{exactly} once, which means one edge going in to the city and one edge
leaving the city, yielding a degree of 2.

Furthermore, when an edge gets added to tour, care must be taken to ensure that
edge does not create a cycle in the graph. This is done by checking if the city
that the edge is going to hasn't already been visited (the degree of the city
is then 0).

\paragraph{Running time}
Let $n$ be the number of cities, then there will $n^2$ edges in the
corresponding graph, because there is an edge between every city.
Inserting an element into a heap takes $O(\log(n))$, where $n$ is the size of
the heap. Since the heap $h$ will have all edges, that is $n^2$, the
\proc{Insert} operation takes $O(\log(n^2)) = O(2\log(n)) = O(\log(n))$.
Therefore, inserting all the edges into the heap will haver running time
$O(n^2\log(n))$. 

The while loop at line \ref{li:greedy:while} in
\proc{Greedy} must in the worst case run $n^2$ times, since it's not certain
that an edge gets added in each round. The body of the while loop performs
only constant-time operations except for \proc{Heap-Extract-Min} which have
running time $O(\log(n^2)) = O(\log(n))$. Therefore, the total running time of
the loop is $O(n^2\log(n))$.

The running time of the last for loop is $O(n)$, since at each round, one city
will be added to the path. This is because in city have one outgoing and one
incoming edge, since they form a tour. The body of the loop will have running
time $O(n)$, since in iterates over all edges. Therefore,
the total running time of the last for loop is $O(n^2)$.

Hence, the total running time of \proc{Greedy} is $O(n^2\log(n))$.

\paragraph{Length of path}
Let $n$ be the number of cities, $opt$ be the optimal path for visiting the $n$
cities and $g$ be the path
found by \proc{Greedy}. The relation between $opt$ and $g$ for any instance $I$ is then
\[\frac{g(I)}{opt(I)} \leq \frac{1}{2}(\lceil \log_2(n) \rceil + 1)\]
according to~\cite{johnson}. 

In the same way as \proc{Nearest-Neighbour} hence, \proc{Greedy} and \proc{Nearest-Neighbour} performs equally bad in the worst case.

\subsection{Optimizing the path}
\label{sec:optimizing_the_path}
The following algorithms are performs \emph{local} optimizations. That is, take an already constructed path,
then make a small local change lowering the local distance and still maintaining a valid path which renders the total length of the path shorter since the total length is just a sum of the local lengths.

\subsubsection{2-opt}
\label{sec:2-opt}
\paragraph{Heuristic}
\begin{codebox}
\Procname{$\proc{Swap}(path, a, b)$}
\li $c \gets \attrib{a}{next}$
\li $d \gets \attrib{b}{next}$
\li $\attrib{c}{prev} \gets d$ 
\li $\attrib{a}{next} \gets b$
\li $\attrib{d}{prev} \gets c$ 
\li $current \gets b$
\li $prev \gets a$
\li \While $current \neq d$
\li     \Do
            $\attrib{current}{next} \gets \attrib{current}{prev}$
\li         $\attrib{current}{prev} \gets \attrib{current}{prev}$
\li         $prev \gets current$
\li         $current \gets \attrib{current}{next}$ 
        \End
\end{codebox}
\begin{codebox}
\Procname{$\proc{2-opt}(path)$}
\li $n \gets \proc{Size}(path)$
\li $\id{improvement} \gets \id{true}$
\li \While $\id{improvement} \isequal \const{true}$
\li     \Do
            $\id{improvement} \gets \const{false}$
\li         $a \gets \proc{First}(path)$
\li         $a_n \gets \attrib{a}{next}$
\li         \For $i \gets 0$ \To $n$
\li             \Do
                    $b \gets \proc{First}(path)$    
\li                 $b_n \gets \attrib{b}{next}$
\li                 \For $j \gets 0$ \To $n$
\li                     \Do
                            \If $a \neq b$ and
                                $a_n \neq b$ and
                                $b_n \neq a$ \label{li:2opt:c1}
\li                             \Then
                                    $e_1 \gets \proc{Distance}(a, a_n) +
                                               \proc{Distance}(b, b_n)$
\li                                 $e_2 \gets \proc{Distance}(a, b) +
                                               \proc{Distance}(a_n, b_n)$
\li                                 \If $e_1 > e_2$ \label{li:2opt:c2}
\li                                     \Then
                                            $\proc{Swap}(path, a, b)$
\li                                         $a_n \gets \attrib{a}{next}$
\li                                         $b_n \gets \attrib{b}{next}$
\li                                         $\id{improvement} \gets
                                             \const{true}$
                                        \End
                                \End
\li                     $b \gets \attrib{b}{next}$
\li                     $b_n \gets \attrib{b_n}{next}$
                        \End
\li             $a \gets \attrib{a}{next}$
\li             $a_n \gets \attrib{a_n}{next}$
                \End
        \End
\end{codebox}

\paragraph{Description}
The idea behind \proc{2-opt} is to traverse the constructed tour and checking
if two edges can be swapped to yield a shorter tour.
\begin{figure}[ht!]
    \centering
    \begin{tikzpicture}
        [ele/.style = {circle, draw, thick, minimum size=6mm,
        font=\small}]
        \node (e2) at (-1, 3) {};
        \node (e3) at (-1, 0) {};
        \node[ele] (e0) at (0,3) {}; 
        \node[ele] (a) at (2, 3) {$a$};
        \node[ele] (c) at (5, 3) {$c$};
        \node[ele] (e1) at (0,0) {}; 
        \node[ele] (b) at (2, 0) {$b$};
        \node[ele] (d) at (5, 0) {$d$};
        \node[ele] (e4) at (7,3) {}; 
        \node[ele] (e5) at (7,0) {}; 
        \node (e6) at (8, 3) {};
        \node (e7) at (8, 0) {};

        \draw[-, loosely dotted, very thick] (e2) -- (e0);
        \draw[-, loosely dotted, very thick] (e3) -- (e1);
        \draw[-, loosely dotted, very thick] (e5) -- (e7);
        \draw[-, loosely dotted, very thick] (e4) -- (e6);
        \draw (e0) -- (a); 
        \draw (e1) -- (b); 
        \draw (a) -- (d); 
        \draw (b) -- (c); 
        \draw (c) -- (e4); 
        \draw (d) -- (e5); 
    \end{tikzpicture}
    \caption{A path before the swap made by \proc{2-opt}}
    \label{fig:2opt:pre}
\end{figure}
\begin{figure}[ht!]
    \centering
    \begin{tikzpicture}
        [ele/.style = {circle, draw, thick, minimum size=6mm,
        font=\small}]
        \node (e2) at (-1, 3) {};
        \node (e3) at (-1, 0) {};
        \node[ele] (e0) at (0,3) {}; 
        \node[ele] (a) at (2, 3) {$a$};
        \node[ele] (c) at (5, 3) {$c$};
        \node[ele] (e1) at (0,0) {}; 
        \node[ele] (b) at (2, 0) {$b$};
        \node[ele] (d) at (5, 0) {$d$};
        \node[ele] (e4) at (7,3) {}; 
        \node[ele] (e5) at (7,0) {}; 
        \node (e6) at (8, 3) {};
        \node (e7) at (8, 0) {};

        \draw[-, loosely dotted, very thick] (e2) -- (e0);
        \draw[-, loosely dotted, very thick] (e3) -- (e1);
        \draw[-, loosely dotted, very thick] (e5) -- (e7);
        \draw[-, loosely dotted, very thick] (e4) -- (e6);
        \draw (e0) -- (a); 
        \draw (e1) -- (b); 
        \draw (a) -- (c); 
        \draw (b) -- (d); 
        \draw (c) -- (e4); 
        \draw (d) -- (e5); 
    \end{tikzpicture}
    \caption{A path after the swap made by \proc{2-opt}}
    \label{fig:2opt:post}
\end{figure}
The idea can be seen in figure \ref{fig:2opt:pre} and figure
\ref{fig:2opt:post}. Therefore, \proc{2-opt} iterates through all possible
quartets ($a$, $a_n$, $b$, $b_n$) of cities and checks if the edges should be
swapped. Any two edges should \emph{not} be swapped in they share a common
node and any two edges should only be swapped if the swap reduces the length of
the path. This is checked by line \ref{li:2opt:c1} and \ref{li:2opt:c1} in
\proc{2-opt}.

Since our implementation of \proc{2-opt} uses a double linked list,
\proc{Swap} needs to ensure that edges between all nodes are correct. When
swapping two edges in double linked list, the list has to be iterated over to 
update all the nodes so their $prev$ and $next$ pointers are consistent.

\paragraph{Running time}
The running time for \proc{Swap} is clearly $O(n)$, since in the worst case,
all edges in the list have to be corrected. The inner for loop in \proc{2-opt}
is run $n$ times, and the body consists only of operations of running time
$O(1)$ and \proc{Swap} operation of running time $O(n)$. Therefore, the total
running time of the inner for loop $O(n^2)$. The outer for loops in being run
$n$ times, and the body of the outer for loop only performs operations with
running time $O(1)$ except for the inner for loop. Therefore, the running time
of the outer for loop is $O(n^3)$.

The question, how many iterations does \proc{2-opt} need before it has
optimized the given path? In the worst case, \proc{2-opt} needs
$\Theta(2^{n/2})$ iterations before finding a local optimal
solution~\cite{johnson}. However, in practice, the number of iterations needed
are usually $\Theta(n)$~\cite{hastad}, which results in total running time of
$O(n^4)$ for this implementation of \proc{2-opt}.

\subsubsection{3-opt}
\label{sec:3-opt}
\paragraph{Heuristic}
\begin{codebox}
\Procname{$\proc{Connect}(a,b)$}
\li     $a.next=b$
\li     $b.prev=a$
\end{codebox}
\begin{codebox}
\Procname{$\proc{3-opt}(path)$}
\li $\id{improvement} \gets \id{true}$
\li \While $\id{improvement} \isequal \const{true} $and$ \id{counter}<\id{N}$
\li     \Do
            $\id{improvement} \gets \const{false}$
\li         $a \gets \proc{First}(path)$
\li         $c \gets a.next.next$
\li         $b \gets c.next.next$
\li         \For ( $;b.next \neq a $ and $ !\id{improvement} $ and $ \id{counter}<\id{N} ; b=b.next $)
\li             \Do
                    $\id{first} \gets \id{true}$
\li                 \For ($;(a\neq \proc{First}(path) $ or $ \id{first} ) $ and $ !better $ and $ counter <N ;a=a.next, b=b.next$)
\li                     \Do
                            $\id{first} \gets \id{false}$
\li                         \For ($ c=a.next.next; c.next\neq b $ and $ !\id{improvement} $ and $ counter < N; c=c.next$)
\li                             \Do
                                    $\id{counter} += 10$
\li                                 $e_0 \gets \proc{Distance}(a, a.next)+
                                               \proc{Distance}(b, b.next)+
                                               \proc{Distance}(c, c.next)$
\li                                 $e_1 \gets \proc{Distance}(a, c.next)+
                                               \proc{Distance}(c, b.next)+
                                               \proc{Distance}(b, a.next)$
\li                                 $e_2 \gets \proc{Distance}(a, b)+
                                               \proc{Distance}(c, b.next)+
                                               \proc{Distance}(a.next, c.next)$

\li                                 \If $e_1<e_0$ and $e_1<e_2$
\li                                     \Then
                                            $an \gets a.next$
\li                                         $\proc{Connect}(a,c.next)$
\li                                         $\proc{Connect}(c,b.next)$
\li                                         $\proc{Connect}(b,an)$
\li                                         $\id{improvement} \gets \id{true}$
\li                                         $counter+=5$
                                        \End
\li                                 \Else \If $e_2 < e_0$
\li                                     \Then
                                            $an=a.next$  
\li                                         $bn=b.next$  
\li                                         $cn=c.next$
\li                                         $current=b.prev$
\li                                         $prev=b$
\li                                         $next=current.prev$ 

\li                                     \While $current\neq c$
\li                                         \Do
                                                $next=current.prev$
\li                                             $\proc{Connect}(prev,current)$
\li                                             $prev=current$
\li                                             $current=next$
\li                                             $counter+=5$
                                            \End
\li                                     $\proc{Connect}(a,b)$
\li                                     $\proc{Connect}(cn,an)$
\li                                     $\proc{Connect}(c,bn)$
\li                                     $\id{improvement} \gets \id{true}$
\li                                     $counter+=10$
                                    \End
                                \End
                            \End
                        \End
                \End
        \End
\end{codebox}

\paragraph{Description}
The idea behind \proc{3-opt} is to traverse the constructed tour in the same way as \proc{2-opt} but instead checking which three edges 
can be swapped to yield a shorter tour.

Therefore, \proc{3-opt} tries to iterate through all possible
tuples ($a$, $a_n$, $c$, $c_n$, $b$, $b_n$) of cities and checks if the edges should be
swapped. The first for-loop sets the width between $a$ and $b$ (starting low to firstly check more local permutations), 
the second for-loop slides $a$ and $b$ around the path and the inner most for-loop iterates $c$ ``inside'' $(a,b)$.
There is 2 distinct path permutations for \proc{3-opt} since all others is either resulting in an illegal path or having already being considered else where in the search, additionaly one needs $c$ ``inside'' $(a,b)$: 
$(a\rightarrow b.next,b\rightarrow c.next,c\rightarrow a.next)$ and $(a\rightarrow c,b\rightarrow c.next,a.next\rightarrow b.next)$.
The one of these (including original) giving the smallest path is switched.
The first permutation doesn't need path-reversal but the second one does between $b$ and $c.next$ and one needs to iterate between those to reverse the double linked list. 

\paragraph{Running time}
All three for-loops are each $O(n)$ since both the max-width, path-length and distance between $a$ and $b$ is each $O(n)$
The first permutation has actually $O(1)$ since it only changes a constant amount of pointers. The second permutation has $O(n)$ from the reversal.
Assuming the number of occurences of both possible improving \proc{3-OPT} permutations is $O(n)$ gives us a resulting complexity for one pass-thru of $O(n^4)$
The question again, how many iterations does \proc{3-opt} need before it has optimized the given path works in the same way as for \proc{2-opt} and the usual (not worstcase) total complexity is $O(n^5)$.

\subsection{Implementation of algorithms}
\label{sec:implementation_of_algorithms}
All heuristics were implemented in the C++ programming language. Of the
heuristics described, \proc{Nearest-Neighbour}, \proc{2-opt} and \proc{3-opt} were
implemented.

\subsubsection{Implementation of \proc{Distance}}
The \proc{Distance} was implemented by calculating all the distances between
all cities and storing the results in a matrix. A call to \proc{Distance} with
arguments $(i, j)$ then simply returned the value of the $i$:th row and the
$j$:th column in the matrix.

\subsubsection{Implementing \proc{Nearest-Neighbour}}
The \proc{Nearest-Neighbour} heuristic was implemented identically to the 
pseudo-code described in section \ref{sec:nearest_neighbour}.

\subsubsection{Implementing \proc{2-opt}}
\label{sec:implementing_2-opt}
The \proc{2-opt} heuristic was also implemented similar to the pseudo-code in
\ref{sec:2-opt}. One detail, which isn't noticeable in the pseudo code, was the
implementation of the double linked list. Since the size of the list is always
known beforehand, all memory that was needed for the list was allocated in the
constructor, instead of allocating a new element of the linked list each time
an element was appended. Then, each insertion of a new element was done 
in-place in the allocated array.

When implementing \proc{2-opt}, we also used a counter to ensure that
\proc{2-opt} didn't run longer than 2 seconds.

\subsubsection{Implementing \proc{3-opt}}
\label{sec:implementing_3-opt}
The \proc{3-opt} was implemented similary to the pseudo-code in \ref{sec:3-opt} with a double linked list. Since the \proc{3-OPT} had such a long running time we implemented a counter that tries to estimate
the number of unit-cost operations and the entire loop is exited once a treshold is reached and as a final touch adjusting this threshold to push as many tries as possible.

